---
title: "Project 3: mypackage Tutorial"
output: rmarkdown::html_vignette
author: "Jessie Chen"
vignette: >
  %\VignetteIndexEntry{mypackage Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(mypackage)
library(tidyverse)
library(tidyr)
library(class)
library(dplyr)
```

## Mypackage Introduction

This package contains useful functions for features such as conducting t-test, 
linear model, predict output with K-nearest neighbors cross validation, and 
random forest cross-validation. The package includes easy to use functions 
that help you in data analysis. 

To install this package, use the code below:
```{r, eval=FALSE}
install.packages("mypackage")
library(mypackage)
```

Alternatively, you can install the development version directly from GitHub:

```{r, eval=FALSE}
devtools::install_github("chen1649/mypackage")
library(mypackage)
```

Now that you have set up the library, let's take a look at some tutorials for 
functions included in mypackage.

### A tutorial for `my_t.test`
First, let's look at the function `my_t.test`:

This function performs a one sample t-test, and returns a list with the the 
numeric test statistic, degrees of freedom, alternative, and p-value. 

To understand how to use this function, you can view the documentation using:

```{r, eval=FALSE}
?my_t.test()
```

Here, we will use the `lifeExp` data from `my_gapminder`.
First, we can set up the `lifeExp` data using:

```{r}
data(my_gapminder)
data_lifeExp <- my_gapminder$lifeExp
```

Now, let's implement this two sided test using this code:

$$
\begin{align}
H_0: \mu &= 60,\\
H_a: \mu &\neq 60.
\end{align}
$$

```{r, eval=FALSE}
my_t.test(data_lifeExp, "two.sided", m = 60)
```

```{r, echo=FALSE}
demo1 <- my_t.test(data_lifeExp, "two.sided", m = 60)
demo1
```

Setting $\alpha$ to 0.05, we ***DO NOT*** reject the null hypothesis because 
the p_value is `r demo1$p_val` which is more than 0.05. 

What about one-sided hypothese? Let's demostrate a test of the hypothesis

$$
\begin{align}
H_0: \mu &= 60,\\
H_a: \mu & < 60.
\end{align}
$$

```{r, eval=FALSE}
my_t.test(data_lifeExp, "less", m = 60)
```

```{r, echo=FALSE}
demo2 <- my_t.test(data_lifeExp, "less", m = 60)
demo2
```

Setting $\alpha$ to 0.05, we reject the null hypothesis because the p_value is 
`r demo2$p_val` which is less than 0.05. 

For the hypothese below, we can use this code:

$$
\begin{align}
H_0: \mu &= 60,\\
H_a: \mu & > 60.
\end{align}
$$

```{r, echo=FALSE}
demo3 <- my_t.test(data_lifeExp, "greater", m = 60)
demo3
```

```{r, eval=FALSE}
my_t.test(data_lifeExp, "greater", m = 60)
```

Setting $\alpha$ to 0.05, we fail to reject the null hypothesis because the 
p_value is `r demo3$p_val` which is more than 0.05. 

### A tutorial for `my_lm`
The function `my_lm` fits a linear model. In the following demonstration,
let's use the lifeExp as response variable and gdpPercap and continent as 
explanatory variables.

```{r}
lm1 <- my_lm(lifeExp ~ gdpPercap + continent, data = my_gapminder)
lm1
```

So what does the result mean? It means that we can expect to see an increase
of  4.452704e-04 in life expectancey for every unit increase of GDP per capita. 
For the standard error, we can say that we expect a difference of 2.349795e-05
in life expectancy if we ran the model again and again. T value measure of how 
many standard deviations our coefficient estimate is far away from 0. In this 
example the t-value is 1.894933e+01 which is quite large compared to the standard
error, indicating a relationship between GDP per capita and life expectancy. 
Lastly, the Pr(>|t|) is the p-value, which means indicates the likelihood of 
observing a relationship by pure chance. In our example, it is 8.552893e-73
which is quite small and we can reject the null hypothesis given that we set 
$\alpha$ to 0.05.

Hypothesis test associated with the the `gdpPerCap` is the following:

$$
\begin{align}
H_0: \mu &= 0,\\
H_a: \mu &\neq 0.
\end{align}
$$

As stated above, we reject the null hypothesis in this case, as we set $\alpha$ 
to 0.05 because we have a p-value of 8.552893e-73.

Now, let's take a look at the actual vs fitted values. 

```{r}
# FINDING fitted yhat
# getting the matrix of the estimates value 
my_estimates <- as.matrix(lm1[,"Estimate"])

# fitting data into matrix to create x matrix
x_mat <- model.matrix(lifeExp ~ gdpPercap + continent, mypackage::my_gapminder)

# matrix multiplication to get yhat 
yhat <- x_mat %*% my_estimates
actual_v_fitted <- data.frame(actual = my_gapminder$lifeExp, fitted = yhat)

# plotting the actual vs fitted
ggplot(actual_v_fitted, aes(x = fitted, y = actual)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, col = "red", lty = 2) + 
  theme_bw(base_size = 15) +
  labs(x = "Fitted values", y = "Actual values", title = "Actual vs. Fitted") +
  theme(plot.title = element_text(hjust = 0.5))
```

So what does this plot say about the model fit? 
The X-axis represents the fitted values, and the Y axis represents the actual 
values. We can see that the points do not fit the linear line closely. The 
linear line represents a overall positive trend, but the actual values are less
linear.


### A tutorial for `my_knn_cv` 

To demonstrate the `my_knn_cv` function, let's predict output class `species` 
using covariates `bill_length_mm`, `bill_depth_mm`, `flipper_length_mm`, and 
`body_mass_g` in the my_penguins data.

```{r}
# setting up the data
data(my_penguins)
data <- na.omit(my_penguins)
# setting train to be bill_length_mm, bill_depth_mm, flipper_length_mm, and 
# body_mass_g
train <- data[3:6] 
cl <- data$species
# create empty list for cv error and train error
cv_err_list <- list(rep(NA, 10))
train_err_list <- list(rep(NA, 10))

# Iterate from k_nn=1,…,10
for (i in 1:10) {
  demo <- my_knn_cv(train, cl, i, 5)
  # record the training misclassification rate and the CV misclassification rate
  cv_err_list[i] <- demo$cv_err
  train_err <- knn(train, train, demo$class)
  train_err_list[i] <- mean(train_err != data$species)
}

# putting results in a table
result <- cbind(cv_err_list, train_err_list)

result
```

According to the table above, the model we to choose is Knn = 1 if we base our 
choice off of training misclassification rate, we would also choose knn = 1
based on the CV misclassification rate. 

In practice, we would choose knn = 1 because we want to choose a knn value 
where the CV misclassification rate is at its lowest. 

At this point, you might wonder why we are using the `my_knn_cv` function in
the first place? The cross-validation is a resampling procedure that uses a 
limited sample in orderto estimate how the model is expected to perform to make
prediction. The process of k-fold cross-validation is as follows:

- Split data into k parts (folds)

- Use all but 1 fold as training data and fit the model

- Use the remaining fold for your test data and make predictions

- Switch which fold is your test data and repeat steps 2 and 3 until all folds
have been test data (k times)

- Compute squared error


### A tutorial for `my_rf_cv`

Let's predict body_mass_g using covariates bill_length_mm, bill_depth_mm, and 
flipper_length_mm.

```{r}
# Create empty lists to store MSE for each k value
MSE_k2 <- rep(NA, 30)
MSE_k5 <- rep(NA, 30)
MSE_k10 <- rep(NA, 30)

# iterates eacch k value 30 times
for (i in 1:30) {
  MSE_k2[i] <-  my_rf_cv(2) 
}

for (i in 1:30) {
  MSE_k5[i] <-  my_rf_cv(5) 
}

for (i in 1:30) {
  MSE_k10[i] <-  my_rf_cv(10) 
}

# append each list together 
cv_list <- append(MSE_k2, MSE_k5)
cv_list <- append(cv_list, MSE_k10)
# create a list of k values
k_val <- c(rep(2, 30), rep(5, 30), rep(10, 30))

# create a dataframe
simulation_csv <- data.frame("k_as_2" = MSE_k2,
                             "k_as_5" = MSE_k5,
                             "k_as_10" = MSE_k10)

# Create a data frame to graph the boxplot
df_rf_cv <- as.data.frame(cbind(cv_list, k_val))
box_plot <- ggplot(data = df_rf_cv, 
                   aes(x = as.factor(k_val), y = as.numeric(cv_list))) +
                   geom_boxplot()

# create summary table
sum_table <- data.frame("k" = c(2, 5, 10),
                      "sd" = c(sd(MSE_k2), sd(MSE_k5), sd(MSE_k10)),
                      "mean" = c(mean(MSE_k2), mean(MSE_k5), mean(MSE_k10)))
```

Below is a boxplot for the MSE errors in each k values, and a summary table
of the standard deviation and mean.

```{r}
box_plot
sum_table
```

The standard deviation and mean for k = 2 are largest and standard deviation and 
mean for k = 10 are smallest. This is because there is a tradeoff 
between bias and variance where a larger k indicates that the algorithm is 
underfitting and a lower k value indicates that it is overfitting. In the case
where k is 2, bias is small but variance is large because it is highly sensitive.
When k is larger, bias is large but variance is small.



