---
title: "Project 3: mypackage Tutorial"
output: rmarkdown::html_vignette
author: "Jessie Chen"
vignette: >
  %\VignetteIndexEntry{mypackage Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(mypackage)
```

## Mypackage Introduction

This package contains useful functions for features such as conducting t-test, linear model, predict output with K-nearest neighbors cross validation, and random forest cross-validation. The package includes easy to use functions that help you in data analysis. 

To install this package, use the code below:
```{r, eval=FALSE}
install.packages("mypackage")
library(mypackage)
```

Alternatively, you can install the development version directly from GitHub:

```{r, eval=FALSE}
devtools::install_github("chen1649/mypackage")
library(mypackage)
```

Now that you have set up the library, let's take a look at some tutorials for functions included in mypackage.

First, let's look at the function `my_t.test`:

This function performs a one sample t-test, and returns a list with the the numeric test statistic, degrees of freedom, alternative, and p-value. 

To understand how to use this function, you can view the documentation using:

```{r, eval=FALSE}
?my_t.test()
```

Here, we will use the `lifeExp` data from `my_gapminder`.
First, we can set up the `lifeExp` data using:

```{r}
data(my_gapminder)
data <- my_gapminder$lifeExp
```

Now, let's implement this two sided test using this code:
$$
\begin{align}
H_0: \mu &= 60,\\
H_a: \mu &\neq 60.
\end{align}
$$

```{r, eval=FALSE}
my_t.test(data, "two.sided", m = 60)
```

```{r, echo=FALSE}
demo1 <- my_t.test(data, "two.sided", m = 60)
demo1
```

Setting $\alpha$ to 0.05, we ***DO NOT*** reject the null hypothesis because the p_value is `r demo1$p_val` which is more than 0.05. 

What about one-sided hypothese? Let's demostrate a test of the hypothesis

$$
\begin{align}
H_0: \mu &= 60,\\
H_a: \mu & < 60.
\end{align}
$$

```{r, eval=FALSE}
my_t.test(data, "less", m = 60)
```

```{r, echo=FALSE}
demo2 <- my_t.test(data, "less", m = 60)
demo2
```

Setting $\alpha$ to 0.05, we reject the null hypothesis because the p_value is `r demo2$p_val` which is less than 0.05. 

For the hypothese below, we can use this code:

$$
\begin{align}
H_0: \mu &= 60,\\
H_a: \mu & < 60.
\end{align}
$$

```{r, echo=FALSE}
demo3 <- my_t.test(data, "less", m = 60)
demo3
```

```{r, eval=FALSE}
my_t.test(data, "greater", m = 60)
```

Setting $\alpha$ to 0.05, we reject the null hypothesis because the p_value is `r demo3$p_val` which is less than 0.05. 
